---
title: "Family Project Tutorial"
author: "[authors]"
date: "TBD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The primary goal of this addendum is to foster familiarity with basic microbial data analysis and figure generation. The data provided to perform this exercise will be used to demonstrate how to create figures in R similar to those found in the main publication.**[more explicit]** In theory, the concepts found in this guide should be applicable to other microbial datasets with some degree of modification.  In order to complete this tutorial basic knowledge of R and its functions is assumed. In this guide concepts are explained at an introductory level supplemented by sample R code for the provided activity. Statistical explanations found in this guide are by no means comprehensive and it is the responsibility of the reader to ensure that the assumptions of any test performed are appropriate for the given data. 

## R Nomenclature 

Here are some tips to help get through RStudio if it’s your first time using it.

1. To see more information on a *function*, such as the inputs taken or output, use:

`help(function)`
or
`?function`

2. Run a line by selecting it and pressing CTRL + ENTER
3. Regions of code that are followed by a # are comments and are not necessary for the code to run. They are to explain what the code is doing e.g.: 
```{r}
df <- mtcars$mpg #assign the mpg column of dataframe mtcars to df
```
4. To complement understanding of particular functions, whenever an existing data frame is modified or a new one is created visualize it either by clicking the data frame in the environment or using the View function e.g.: `View(df)`

## Preparing Data for Figures 

 To create figures it is important to ensure that the data is prepared and formatted correctly. Before beginning, specify a desired working directory in R-studio. The metadata, taxonomy, and OTU tables found at **(location we are making them accessible)** are the only files needed for the completion of this tutorial. 
 
### Data Import 

First set the appropriate working directory, i.e. where we want R to look for our files. Change `File Path` to the location where your desired data is being kept. *It's my file path for now but when were done change back to "File Path".*
```{r}
setwd("E:/School/ICU/Bio/199/Family project - Quarter 1/Family Project/Figs Practice/Attempting FP figs from ABC")
```

Import the metadata file, taxonomy, and OTU Table. For the taxonomy and OTU table the first column can be specified as the row names, the metadata file will require some extra steps to be formatted properly. 
```{r}
OTU.taxonomyFP <- read.delim("taxonomy.txt", row.names = 1)
OTU.tableFP <- read.delim("OTU_table.txt", row.names = 1)
metadataFP <- read.delim("Microbiome_mapping_14Oct2019.txt")
```

Ensure that all the desired files maintain a consistent formatting style. For the metadata data frame the row names must be formatted for consistency across the other files. The row names must be in the format of “X.##” -  “##” being indicative of the sample number. 
```{r}
metadataFP$X.NAME <- interaction( "X", metadataFP$X.NAME, sep = "")
rownames(metadataFP) <- metadataFP[,1]
metadataFP[,1] <- NULL
```

### Data Filtering 

It is important to determine if filtering is needed before continuing. An example utilizing the sample data is provided below, however, note that this procedure will not be implemented when creating the final figures at later stages in this tutorial. 

Merge the OTU table with taxonomy into a single data frame. 
```{r}
OTU.table.with.taxonomyFP <- as.data.frame(merge(OTU.taxonomyFP, 
                                               OTU.tableFP, by.x = "row.names", by.y = "row.names"))
```

Remove unassigned taxa from the merged data frame. For certain diversity metrics filtering may be undesired **citation needed**
```{r}
OTU.table.filteredtaxaFP <- OTU.table.with.taxonomyFP[!grepl("Unassigned", 
                                                        OTU.table.with.taxonomyFP$Kingdom),]
```

Subset the original OTU table with the filtered taxa to create a filtered OTU table. 
```{r}
FPfilteredtaxa <- OTU.table.filteredtaxaFP$Row.names
FPfilterOTU <- subset(OTU.tableFP, rownames(OTU.tableFP) %in% FPfilteredtaxa)
```
















